### 2022-03-21 outputs/paper/paper.Rmd
figures updated

---
title: "My title"
author: 
  - Eunkyo Oh
thanks: "Code and data are available at: https://github.com/oheunkyo/."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Introduction

# Data

The data we are going to analyze is the result of Canadian General Social Survey on Family (2017). The data is organized in respect to its age, sex, education, number of children, and other variables totaling up to eighty one entries. This survey was performed from February 2 to November 30, 2017, and it may not correctly represent the current state of the respondents' familial status. Each respondent is fifteen years old or older, a full-time residents in Canada excluding Yukon, Northwest Territories, and Yukon. There are 20602 responses in the data. We obtain our dataset from the University of Toronto Library, available to University of Toronto students and faculty members. This data can also be found in databases of other academic institutions to their staff and students. We modify and clean our data using packages `tidyverse` [@citetidyverse] and `janitor` [@citejanitor] using the  and the statistical programming language R [@citeR].

How we organized the data

Table of dataset sneak peek

```{r dataextract, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(dplyr)

# Read in the csv
gssurvey <- read_csv(here::here("outputs/data/gss.csv"))

# Pick only the necessary columns
gssurvey <- 
  gssurvey |>
  select(age, 
         feelings_life, 
         sex,
         self_rated_health, 
         self_rated_mental_health, 
         income_family)

gssurveyagegroup <- 
  gssurvey |> 
  mutate(age_group = 
           case_when(age < 26 ~ "16-25",
                     age < 36 ~ "26-35",
                     age < 46 ~ "36-45",
                     age < 56 ~ "46-55",
                     age < 66 ~ "56-65",
                     age < 76 ~ "66-75",
                     age < 100 ~ ">76"
                     ),
         age_group = factor(age_group,
                            levels = c("16-25",
                                       "26-35",
                                       "36-45",
                                       "46-55",
                                       "56-65",
                                       "66-75",
                                       ">76"
                                       )
                            )
         )

gssurvey |> 
  slice(1:10) |> 
  kable(
    caption = "sdsd",
    #col.names = c("Municipality", "Name of Location", "Religion of Practice"),
    #digits = 1,
    booktabs = TRUE, 
    linesep = ""
  )

```

```{r feelingscount, fig.cap="Number of Perceived Feelings on Life Responses", echo = FALSE, warning=FALSE, message=FALSE}
gssurvey |> 
  ggplot(mapping = aes(x = feelings_life)) +
  geom_histogram(bins = 11, fill="aquamarine4") +
  theme_minimal() +
  labs(x = "Feelings on life (ten point scale)",
       y = "Number of respondents") +
  scale_x_continuous(breaks=seq(0, 10, 1))
```

```{r feelingsvssex, fig.cap="Number of Perceived Feelings on Life Responses by Sex", echo = FALSE, warning=FALSE, message=FALSE}
gssurvey |> 
  ggplot(mapping = aes(x = feelings_life, fill = sex)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(x = "Feelings on life (ten point scale)",
       y = "Number of respondents",
       fill = "Sex") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_continuous(breaks=seq(0, 10, 1))
```

```{r agecount, fig.cap="Number of Responses by Age Group", echo = FALSE, warning=FALSE, message=FALSE}
gssurveyagegroup |>  
  ggplot(mapping = aes(x = age_group, fill = feelings_life)) +
  geom_bar(fill="aquamarine4") +
  theme_minimal() + 
  labs(x = "Age groups",
       y = "Number of respondents")
```

```{r averagebyage, echo=FALSE, warning=FALSE, message=FALSE}
# making a new data to find proportions of people per location
summary_gss <- 
  gssurveyagegroup |> 
    group_by(age_group) |> 
    summarise(num = n(), average = mean(feelings_life, na.rm = TRUE))

# Creating a table 
summary_gss |>
    kable(
      caption = "Mean Feeling of Life by Age Group",
      col.names = c("Age Group", 
                    "Number of Respondents", 
                    "Average Feeling of Life"),
      digits = 1,
      booktabs = TRUE, 
      linesep = ""
    )
```

```{r incomevsfeelings, fig.cap="Number of Perceived Feelings on Life Responses by Household Income", echo = FALSE, warning=FALSE, message=FALSE}
level_order <- c('Less than $25,000', '$25,000 to $49,999', 
                 '$50,000 to $74,999', '$75,000 to $99,999', 
                 '$100,000 to $124,999',
                 '$125,000 and more')

gssurvey |>
  filter(income_family %in% level_order) |>
  ggplot(mapping = aes(x = factor(income_family, levels = level_order), y = feelings_life, color = income_family)) +
  geom_jitter() +
  geom_smooth() +
  theme_minimal() +
  labs(x = "Household Income (CAD)",
       y = "Feelings on life (ten point scale)") +
  guides(color="none") +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(breaks=seq(0, 10, 1))

```

```{r healthvsfeelings, fig.cap="Self Rated Health versus Feeling on Life and Self Rated Mental Health versus Feeling on Life",echo=FALSE, warning=FALSE, message=FALSE}

library(patchwork)

level_order <- c('Poor', 'Fair', 'Good', 'Very good', 'Excellent')
health <-
  gssurvey |>
  filter(self_rated_health %in% level_order) |>
    ggplot(mapping = aes(x = factor(self_rated_health, level=level_order), 
                         y = feelings_life, color = self_rated_health)) +
    geom_jitter() +
    geom_smooth() +
    theme_minimal() +
    labs(x = "Self rated health (ten point scale)",
         y = "Feelings on life (ten point scale)",
         caption = "Figure 5: Self Rated \n Health versus Feeling on Life") +
    guides(color="none") +
    scale_fill_brewer(palette = "Dark2") +
    scale_y_continuous(breaks=seq(0, 10, 1))

mhealth <-
  gssurvey |>
  filter(self_rated_mental_health %in% level_order) |>
    ggplot(mapping = aes(x = factor(self_rated_mental_health, level=level_order), 
                         y = feelings_life, color = self_rated_mental_health)) +
    geom_jitter() +
    geom_smooth() +
    theme_minimal() +
    labs(x = "Self rated mental health (ten point scale)",
         y = "Feelings on life (ten point scale)",
         caption = "Figure 6: Self Rated Mental \n Health versus Feeling on Life") +
    guides(color="none") +
    scale_fill_brewer(palette = "Dark2") +
    scale_y_continuous(breaks=seq(0, 10, 1))

(health + mhealth)
```
removeed NA and DOn't KNow responses

# Results

# Discussion

## First discussion point

why ____ has such high count

## Second discussion point

why ____ has such high correlation

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References